High-order cognition is supported by complex but compressible brain activity patterns

### 2023-02-22 paper/main.tex
working on intro, pie figure, added information and compressibility figure

We take as a given that brain activity is highly flexible: our brains can
exhibit nearly infinite activity patterns. However, those activity patterns are
also highly structured. For example, full-brain correlation matrices are stable
within~\citep{FinnEtal15} and across~\citep{YeoEtal11} individuals. This
stability implies that our brains are not

For example, despite stable individual differences in
full-brain correlations~\citep{FinnEtal15},

A number of prior studies suggest that

@@

Closely related to the above notions of expressiveness versus robustness are
measures of how much \textit{information} is contained in a given signal or
pattern, and how \textit{redundant} a signal is. Formally, information is
defined as the amount of uncertainty about a given variables' outcomes (i.e.,
entropy), measured in \textit{bits}, or the optimal number of yes/no questions
needed to reduce uncertainty about the variable's outcomes to zero. Highly
complex systems with many degrees of freedom (i.e., high flexibility and
expressiveness), are more information-rich than simpler or more constrained
systems. The redundancy of a signal denotes the difference how expressive the
signal \textit{could} be (i.e., proportional to the number of unique states or
symbols used to transmit the signal) and the actual information rate (i.e., the
entropy of each individual state or symbol). If a brain network's nodes are
fully independent, then the number of bits required to express a single
activity pattern is proportional to the number of nodes. The network would also
be minimally redundant, since the status of every node would be needed to fully
express a single brain activity pattern. If a brain network's nodes are fully
coupled and identical, then the number of bits required to express a single
activity pattern is proportional to the number of unique states or values any
individual node can take on. Such a network would be highly redundant, since
knowing any individual node's state would be sufficient to recover the
full-brain activity pattern. Highly redundant systems are also robust, since
there is little information loss from losing any given observation.

% Some papers to add in:

% https://www.nature.com/articles/s41467-019-13930-8
% https://onlinelibrary.wiley.com/doi/full/10.1002/jnr.24316?casa_token=PjnffsBh3vgAAAAA%3AynPc_easF6EXPrEK7ihnQ6Jc8QeWq6x0u-ytDsMwxQ4a0jnIvXhg8xBQDDVnAoMbYhUWiwUtlj4vTA
% https://www.sciencedirect.com/science/article/pii/S1053811918305238?casa_token=xSH5EIN1rZYAAAAA:R2Zf9Lu_np3v0t5-lNZNal0JaJLk1JyYe4QxtmHnba9_lMGJUJgKNK4lHoybuITemQqYUIcuNA
% https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3174820/
% https://www.sciencedirect.com/science/article/pii/S0896627314004000
% https://www.pnas.org/doi/abs/10.1073/pnas.0905267106
% https://www.sciencedirect.com/science/article/pii/S0896627318302411
% https://www.liebertpub.com/doi/full/10.1089/brain.2011.0068
% https://www.sciencedirect.com/science/article/pii/S1053811917302872#f0015
% https://www.sciencedirect.com/science/article/abs/pii/S1053811919303027
% https://pubmed.ncbi.nlm.nih.gov/34625348/

If brain activity patterns contain rich task-relevant information, we should be
able to use the activity patterns to accurately differentiate between different
aspects of the task~\citep[e.g., using pattern classifiers;][]{NormEtal06b}.
For example, prior work has shown a direct correspondence between
classification accuracy and the information content of a signal~\citep{Alva02}.
If brain activity patterns are

>>>>

@@

Closely related to the above notions of expressiveness versus robustness are
measures of how much \textit{information} is contained in a given signal or
pattern, and how \textit{redundant} a signal is~\citep{Shan48}. Formally,
information is defined as the amount of uncertainty about a given variables'
outcomes (i.e., entropy), measured in \textit{bits}, or the optimal number of
yes/no questions needed to reduce uncertainty about the variable's outcomes to
zero. Highly complex systems with many degrees of freedom (i.e., high
flexibility and expressiveness), are more information-rich than simpler or more
constrained systems. The redundancy of a signal denotes the difference how
expressive the signal \textit{could} be (i.e., proportional to the number of
unique states or symbols used to transmit the signal) and the actual
information rate (i.e., the entropy of each individual state or symbol). If a
brain network's nodes are fully independent, then the number of bits required
to express a single activity pattern is proportional to the number of nodes.
The network would also be minimally redundant, since the status of every node
would be needed to fully express a single brain activity pattern. If a brain
network's nodes are fully coupled and identical, then the number of bits
required to express a single activity pattern is proportional to the number of
unique states or values any individual node can take on. Such a network would
be highly redundant, since knowing any individual node's state would be
sufficient to recover the full-brain activity pattern. Highly redundant systems
are also robust, since there is little information loss from losing any given
observation.

We take as a given that brain activity is highly flexible: our brains can
exhibit nearly infinite activity patterns. This flexibility implies that our
brains activity patterns are highly information rich. However, brain activity
patterns are also highly structured. For example, full-brain correlation
matrices are stable within~\citep{FinnEtal15, FinnEtal17, GratEtal18} and
across~\citep{YeoEtal11, GlerEtal12, GratEtal18, ColeEtal14} individuals. This
stability suggests that our brains' activity patterns are at least partially
constrained, for example by anatomical, external, or internal factors.
Constraints on brain activity that limit its flexibility decrease
expressiveness (i.e., its information rate). However, constraints on brain
activity also increase its robustness to noise (e.g., ``missing'' or corrupted
signals may be partially recovered). For example, recent work has shown that
full-brain activity patterns may be reliably recovered from only a relatively
small number of implanted electrodes~\citep{OwenEtal20, ScanEtal21}. This
robustness property suggests that the relevant signal (e.g., underlying factors
that have some influence over brain activity patterns) are compressible.

To the extent that brain activity patterns contain rich task-relevant
information, we should be able to use the activity patterns to accurately
differentiate between different aspects of the task~\citep[e.g., using pattern
classifiers;][]{NormEtal06b}. For example, prior work has shown a direct
correspondence between classification accuracy and the information content of a
signal~\citep{Alva02}. To the extent that brain activity patterns are
compressible, we should be able to generate simplified (e.g., lower
dimensional) representations of the data while still preserving the relevant or
important aspects of the original signal. In general, information content and
compressibility are related but are partially dissociable
(Fig.~\ref{fig:information--compression}).

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/pie_example.pdf}
  \caption{\textbf{Illustration of compression.} Visual analogy for
    neural compression. Here are 2 images of pies, one more complex
    than the other. \textbf{a.} It takes fewer components to reach the same percent
    variance explained in the less complex pie, which corresponds to
    higher compression. \textbf{b.} However, with very few components, similar variance is explained in both pies.
\textbf{c.} Plots the cumulative explained variance for more and more components.}
  \label{fig:information-compression}
\end{figure}

Specifically, if a given signal
(e.g., a representation of brain activity patterns) contains more information
about ongoing cognitive processes, then the peak decoding accuracy should be
high. And if the signal is compressible, a low-dimensional embedding of the
signal will be similarly informative to the original signal. A signal can be
informative but not compressible, informative but
